% report/report.tex
\documentclass[11pt,a4paper]{article}

% --------- Packages ---------
\usepackage[margin=2.5cm]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{listings}

\hypersetup{
  colorlinks=true,
  linkcolor=blue,
  urlcolor=blue,
  citecolor=blue
}

% --------- Listings (AnB / OFMC outputs) ---------
\lstdefinestyle{anb}{
  basicstyle=\ttfamily\small,
  breaklines=true,
  frame=single,
  columns=fullflexible,
  showstringspaces=false
}

\lstdefinestyle{terminal}{
  basicstyle=\ttfamily\small,
  breaklines=true,
  frame=single,
  columns=fullflexible,
  showstringspaces=false
}

% --------- Simple helper command ---------
\newcommand{\sectionauthor}[1]{\noindent\textbf{Section author: #1}\par\medskip}

% --------- Title info ---------
\title{Open Authorization: Protocol Design and Formal Analysis with OFMC}
\author{
  Kristófer Birgir Hjörleifsson - s253155 \and Mikael Máni Eyfeld Clarke - s253024
}
\date{Hand-in: Mar 16, 2026}

\begin{document}
\maketitle

\begin{abstract}
We design and incrementally refine an open authorization protocol in which a user A delegates
read-only access to a photo resource at server B to a photo book service P, using an honest identity
provider \texttt{idp} (modeled as a lowercase constant to prevent the intruder from impersonating it).
The protocol is specified in AnB and analyzed using OFMC under the Dolev--Yao attacker
model. This report documents the development process, discovered attacks, design iterations, and the
final protocol, including weekly special tasks.
\end{abstract}

\tableofcontents
\newpage

% ------------------------------------------------
\section{Group Setup and Section Ownership}
\sectionauthor{Kristófer Birgir Hjörleifsson}

\begin{itemize}[leftmargin=*]
  \item \textbf{Group members:} Kristófer Birgir Hjörleifsson, Mikael Máni Eyfeld Clarke
  \item \textbf{Repository:} (https://github.com/kristoferbirgir/Project1-LogicForSecurity)
  \item \textbf{Submitted artifacts:} This report (PDF), AnB files, and the lab logbook.
\end{itemize}

\noindent\textbf{Section ownership requirement (DTU):} Each section in this report lists exactly one
author. Any section without a single designated author may be counted as not submitted.

% ------------------------------------------------
\section{Resources Used}
\sectionauthor{Kristófer Birgir Hjörleifsson}

\begin{itemize}[leftmargin=*]
  \item Course slides and lecture notes (weeks 2--7).
  \item OFMC / AnB documentation and examples from the course material.
  \item AVISPA project documentation for AnB syntax reference.
  \item AI tools: GitHub Copilot, Claude (Anthropic), and ChatGPT (OpenAI) used for protocol development iteration and documentation.
  \item Group discussions between team members.
\end{itemize}

% ------------------------------------------------
\section{Scenario and Security Goals}
\sectionauthor{Mikael Máni Eyfeld Clarke}

\subsection{Scenario}
A stores holiday photos at server B. A wants to create a photo book using service P and wishes to
authorize P to access a subset of A's photos at B without downloading and re-uploading them.
A authenticates using an identity provider \texttt{idp} via a shared password \texttt{pw(A,idp)}.
The password is considered guessable and must not be used as an encryption key.
All parties except A have public/private keys.
All parties know \texttt{idp}'s public key, and \texttt{idp} knows all public keys.
In the AnB model, \texttt{idp} is written in lowercase to declare it as an honest constant---the
intruder cannot instantiate this role.

\subsection{Agents}
\begin{itemize}[leftmargin=*]
  \item \textbf{A (User):} Owns photos, authenticates via password, no cryptographic keys.
  \item \textbf{B (Photo Server):} Stores A's photos, has key pair \texttt{pk(B)/inv(pk(B))}.
  \item \textbf{P (Printing Service):} Wants to access A's photos, has key pair \texttt{pk(P)/inv(pk(P))}.
  \item \textbf{idp (Identity Provider):} Authenticates A, issues authorization tokens, has key pair \texttt{pk(idp)/inv(pk(idp))}. Modeled as a lowercase constant (honest, trusted).
\end{itemize}

\subsection{Security and Functional Goals}
We target the following goals:
\begin{itemize}[leftmargin=*]
  \item \textbf{Authorization:} B should only release photos if authorized by A via \texttt{idp}.
  \item \textbf{Password secrecy:} A's password \texttt{pw(A,idp)} must not be revealed to P, B, or the intruder.
  \item \textbf{Photo confidentiality:} Photos should remain secret between B and P.
  \item \textbf{Authentication:} B must be able to verify \texttt{idp} signed the authorization token.
\end{itemize}

% ------------------------------------------------
\section{Week 2: Informal Design and Static Dolev--Yao Analysis}
\sectionauthor{Kristófer Birgir Hjörleifsson}

\subsection{Informal Outline}
The basic protocol flow:
\begin{enumerate}
  \item A requests authorization from \texttt{idp} (encrypted with \texttt{idp}'s public key)
  \item \texttt{idp} creates signed token authorizing P to access A's photos at B
  \item A forwards token to P
  \item P presents token to B (encrypted with B's public key)
  \item B verifies token and sends photos to P (encrypted with P's public key)
\end{enumerate}

\subsection{First AnB Model (Simplified)}
Initial simplification: A has a key pair (temporary, removed in Week 3).
\texttt{idp} is lowercase---an honest constant that the intruder cannot instantiate.

\begin{lstlisting}[style=anb,caption={Week 2 protocol}]
Actions:
  A -> idp: {A, P, B}(pk(idp))
  idp -> A: {A, P, B}(inv(pk(idp)))
  A -> P: {A, P, B}(inv(pk(idp)))
  P -> B: {{A, P, B}(inv(pk(idp)))}(pk(B))
  B -> P: {photos(A)}(pk(P))
\end{lstlisting}

\subsection{OFMC Result}
OFMC found an attack on the secrecy goal: the intruder plays A's role, requests authorization
for itself, and manipulates the token flow so that B sends photos encrypted to a key the intruder
controls. This is possible because A, B, P are all uppercase variables in AnB, meaning the
intruder can instantiate them.

\textbf{Key insight:} Because \texttt{idp} is lowercase, the intruder \emph{cannot} forge
\texttt{idp}'s signature. The attack is about role manipulation, not signature forgery.

\subsection{Special Task: Static Analysis (Passive Intruder)}
\textbf{Assumption:} Intruder only eavesdrops (passive), does not inject messages.

\textbf{What intruder observes:}
\begin{itemize}[leftmargin=*]
  \item Message 1: \texttt{\{A, P, B\}(pk(idp))} -- cannot decrypt (no \texttt{inv(pk(idp))})
  \item Message 2: \texttt{\{A, P, B\}(inv(pk(idp)))} -- can verify signature, learns A, P, B
  \item Message 3: Same signed token -- learns nothing new
  \item Message 4: \texttt{\{\{token\}\}(pk(B))} -- cannot decrypt outer layer
  \item Message 5: \texttt{\{photos(A)\}(pk(P))} -- cannot decrypt
\end{itemize}

\textbf{Intruder's knowledge (passive):} Agent identities A, P, B (from signed token). \textbf{Cannot learn:} photos, password, or private keys.

\textbf{Conclusion:} Protocol is secure against passive eavesdropping. The active attacks require message injection and role manipulation.

% ------------------------------------------------
\section{Week 3: Identity Provider and Lazy Intruder Executability}
\sectionauthor{Mikael Máni Eyfeld Clarke}

\subsection{Protocol Refinement}
\textbf{Key change:} A no longer has a key pair. A authenticates to \texttt{idp} using
shared password \texttt{pw(A,idp)}.

\begin{lstlisting}[style=anb,caption={Week 3: Password authentication}]
A -> idp: {A, P, B, pw(A,idp)}(pk(idp))
\end{lstlisting}

The password is encrypted under \texttt{idp}'s public key, so the intruder cannot learn it.
Password is NOT used as an encryption key---only as a credential.

\subsection{Special Task: Lazy Intruder Method (P Executable by Intruder)}
We show role P is executable by the intruder using constraint solving (lazy intruder).

\textbf{Role P's messages:}
\begin{itemize}[leftmargin=*]
  \item Receives from A: \texttt{token = \{A, P, B\}(inv(pk(idp)))}
  \item Sends to B: \texttt{\{\{A, P, B\}(inv(pk(idp)))\}(pk(B))}
\end{itemize}

\textbf{Constraint solving (lecture style):}

\noindent Initial intruder knowledge $\mathit{IK}_0$: all public names, \texttt{pk(B)}, \texttt{pk(P)}, \texttt{pk(idp)}, agent names.

\noindent After intercepting message 3 (A $\to$ P):
\[
\mathit{IK}_1 = \mathit{IK}_0 \cup \{\texttt{\{A, P, B\}(inv(pk(idp)))}\}
\]

\noindent The intruder must compose:
\texttt{\{\{A, P, B\}(inv(pk(idp)))\}(pk(B))}

\noindent Resolution steps:
\begin{enumerate}
  \item \textbf{Axiom:} \texttt{\{A, P, B\}(inv(pk(idp)))} $\in \mathit{IK}_1$ \quad (intercepted token)
  \item \textbf{Axiom:} \texttt{pk(B)} $\in \mathit{IK}_0$ \quad (public key)
  \item \textbf{Compose (encryption):} From steps 1--2, compose
        \texttt{\{\{A, P, B\}(inv(pk(idp)))\}(pk(B))}
\end{enumerate}

\textbf{Conclusion:} Role P is executable by the intruder. All outgoing messages can be
constructed from the intruder's knowledge via Axiom and Compose rules.

% ------------------------------------------------
\section{Week 4: Typing, Formats, and Public-Key Lookup}
\sectionauthor{Kristófer Birgir Hjörleifsson}

\subsection{Type-Flaw Resistance via Format Tags}
Type flaws occur when an intruder substitutes a message of one type where another is expected.
We prevent this by adding constant format tags (\texttt{f1}--\texttt{f4}) as the first element of
each message tuple, making every message structurally unique:

\begin{lstlisting}[style=anb,caption={Week 4: Format-tagged messages}]
Msg 1: {f1, A, P, B, pw(A,idp)}(pk(idp))      % Auth request
Msg 2: {f2, A, P, B}(inv(pk(idp)))            % Token
Msg 4: {f3, P, {f2, A, P, B}(inv(pk(idp)))}(pk(B))  % Photo request
Msg 5: {f4, B, photos(A)}(pk(P))              % Photo response
\end{lstlisting}

The tag \texttt{f1} marks auth requests, \texttt{f2} marks tokens, etc. Since each message
begins with a distinct constant, no two different message types can be unified.

\subsection{Public-Key Lookup Subprotocol}
A separate protocol allows A to obtain public keys from \texttt{idp}:

\begin{lstlisting}[style=anb,caption={Key lookup protocol}]
A -> idp: {f5, A, B, pw(A,idp)}(pk(idp))
idp -> A: {f5, A, B, pk(B)}(inv(pk(idp)))
\end{lstlisting}

\subsection{Special Task: Type-Flaw Resistance Proof}
\textbf{Definition:} A protocol is type-flaw resistant if for any two non-variable messages $M_1$, $M_2$ in the set of sub-messages of the protocol (SMP), if they have a unifier $\sigma$, then
$\text{type}(M_1) = \text{type}(M_2)$.

\textbf{Analysis of message types:}
\begin{center}
\begin{tabular}{lll}
\toprule
Message & Leading tag & Arity \\
\midrule
Msg 1 & f1 & 5 (f1, Agent, Agent, Agent, Password) \\
Msg 2 & f2 & 4 (f2, Agent, Agent, Agent) \\
Msg 4 outer & f3 & 3 (f3, Agent, SignedToken) \\
Msg 5 & f4 & 3 (f4, Agent, Data) \\
\bottomrule
\end{tabular}
\end{center}

\textbf{Unification check:}
\begin{itemize}[leftmargin=*]
  \item No two messages share the same leading tag, so no unifier exists between different message types.
  \item Even ignoring tags: Msg 1 has 5 components, Msg 2 has 4, etc.---different arities prevent unification.
  \item The signed token in message~2 and the inner token carried inside message~4 do unify,
        but this is benign because they are the \emph{same} tagged message type~\texttt{f2}.
        Type-flaw resistance only requires that messages of \emph{different} types cannot unify.
\end{itemize}

\textbf{Conclusion:} Protocol is type-flaw resistant.

% ------------------------------------------------
\section{Week 5: Channels (TLS-like Pseudonymous Channels)}
\sectionauthor{Mikael Máni Eyfeld Clarke}

\subsection{Channel Notation}
\begin{center}
\begin{tabular}{ll}
\toprule
Notation & Meaning \\
\midrule
\texttt{A -> B: M} & Insecure (Dolev-Yao) \\
\texttt{A *-> B: M} & Confidential only \\
\texttt{A ->* B: M} & Authenticated only \\
\texttt{A *->* B: M} & Secure channel (mutual TLS) \\
\texttt{[A] *->* B: M} & A pseudonymous, B authenticated (TLS server-auth) \\
\texttt{B *->* [A]: M} & Reply to pseudonymous A \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Pseudonymous Channel Variant}
\begin{lstlisting}[style=anb,caption={Week 5: Pseudonymous TLS channels}]
[A] *->* idp: f1, A, P, B, pw(A,idp)
idp *->* [A]: {f2, A, P, B}(inv(pk(idp)))
[A] *->* P: {f2, A, P, B}(inv(pk(idp)))
P *->* B: f3, P, {f2, A, P, B}(inv(pk(idp)))
B *->* P: f4, B, photos(A)
\end{lstlisting}

The \texttt{[A]} notation means A is \emph{pseudonymous}: the server authenticates itself (has
a certificate) but does not know the client's identity. This models TLS with server-side
authentication only. A authenticates at the application level via \texttt{pw(A,idp)}.

\subsection{Special Task: What Was Replaced}
\begin{center}
\begin{tabular}{lll}
\toprule
Original & Channel Version & Replaced? \\
\midrule
\texttt{\{M\}(pk(idp))} & \texttt{[A] *->* idp: M} & Yes \\
\texttt{\{M\}(pk(B))} & \texttt{P *->* B: M} & Yes \\
\texttt{\{M\}(pk(P))} & \texttt{B *->* P: M} & Yes \\
\texttt{\{M\}(inv(pk(idp)))} & Cannot replace & \textbf{No} \\
\bottomrule
\end{tabular}
\end{center}

\textbf{Why the signature cannot be replaced:} B needs to verify \texttt{idp}'s approval, but B
has no direct TLS channel to \texttt{idp}. The token travels through A and P, so B must verify
the signature cryptographically.

% ------------------------------------------------
\section{Week 6: Privacy and Guessable Password}
\sectionauthor{Kristófer Birgir Hjörleifsson}

\subsection{Guessable Secrets}
A guessable secret (like a password) can be subject to offline dictionary attacks. The intruder
captures data containing or derived from the password and tries verifying guesses offline.

\subsection{Security with Guessable Password}
Our protocol protects against guessing attacks because:
\begin{itemize}[leftmargin=*]
  \item Password is sent over a pseudonymous secure channel (\texttt{[A] *->* idp})
  \item Intruder never sees any derivation of the password on the wire
  \item No offline verification possible
\end{itemize}

\textbf{OFMC Goal:} \texttt{pw(A,idp) guessable secret between A, idp}

\textbf{Result:} No guessing attack found. The only attack is on photo secrecy (intruder plays A),
not on the password.

\subsection{Special Task: Insecure Variant Demonstration}
We created an insecure variant using hash-based challenge-response over an insecure channel.
This is a proper example of offline guessing vulnerability (unlike public-key encryption,
which does \emph{not} allow the intruder to verify guesses):

\begin{lstlisting}[style=anb,caption={Insecure version: hash-based challenge-response}]
A -> idp: A, P, B
idp -> A: NB
A -> idp: h(pw(A,idp), NB)
\end{lstlisting}

The intruder observes \texttt{NB} and \texttt{h(pw(A,idp), NB)} on the wire and can
verify password guesses by recomputing the hash.

\textbf{OFMC Result:} \texttt{ATTACK\_FOUND (guesswhat)}
\begin{lstlisting}[style=terminal,caption={Guessing attack trace}]
(A,1) -> i: A, P, B
i -> (A,1): NB
(A,1) -> i: h(pw(A,idp), NB)
i can produce secret h(guessPW, NB)
\end{lstlisting}

\textbf{Conclusion:} Public-key encryption (\texttt{\{pw\}(pk(idp))}) does \emph{not} allow
offline guessing because the intruder cannot decrypt without \texttt{inv(pk(idp))}.
Hash-based challenge-response \emph{does} allow it because the intruder can recompute
\texttt{h(guess, NB)} and compare. Secure channels prevent both.

% ------------------------------------------------
\section{Final Protocol (AnB Specification Summary)}
\sectionauthor{Mikael Máni Eyfeld Clarke}

\subsection{Final Message Flow}
\begin{lstlisting}[style=anb,caption={Final protocol (photo\_auth\_final.AnB)}]
[A] *->* idp: f1, A, P, B, pw(A,idp)
idp *->* [A]: {f2, A, P, B}(inv(pk(idp)))
[A] *->* P: {f2, A, P, B}(inv(pk(idp)))
P *->* B: f3, P, {f2, A, P, B}(inv(pk(idp)))
B *->* P: f4, B, photos(A)
\end{lstlisting}

\textbf{Step-by-step:}
\begin{enumerate}
  \item A authenticates to \texttt{idp} with password over pseudonymous TLS
  \item \texttt{idp} creates signed authorization token (format tag \texttt{f2})
  \item A forwards token to P over pseudonymous TLS
  \item P presents token to B with mutual TLS (format tag \texttt{f3})
  \item B verifies token signature and sends photos to P (format tag \texttt{f4})
\end{enumerate}

\subsection{Initial Knowledge}
\begin{center}
\begin{tabular}{ll}
\toprule
Agent & Knowledge \\
\midrule
A & A, B, P, idp, pw(A,idp), f1, f2, f3, f4 \\
B & A, B, P, idp, pk(B), pk(idp), inv(pk(B)), photos(A), f1, f2, f3, f4 \\
P & A, B, P, idp, pk(B), pk(P), pk(idp), inv(pk(P)), f1, f2, f3, f4 \\
idp & A, B, P, idp, pk(B), pk(P), pk(idp), inv(pk(idp)), pw(A,idp), f1, f2, f3, f4 \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Goals}
\begin{itemize}[leftmargin=*]
  \item \texttt{B authenticates idp on f2, A, P, B} -- B verifies \texttt{idp} signed the token
  \item \texttt{photos(A) secret between B, P} -- Photos only known to B and P
  \item \texttt{pw(A,idp) guessable secret between A, idp} -- Password protected
\end{itemize}

\subsection{Verification Results}
\textbf{OFMC Result:} ATTACK\_FOUND (secrets)

OFMC reports a secrecy violation because the intruder can instantiate the user role~A with itself
(A is an uppercase variable). In that case, \texttt{idp} legitimately issues a token for the
intruder's own identity, and the intruder eventually learns \texttt{photos(i)}---i.e.\ its own
resource. This is not a meaningful break of the intended authorization property for honest users;
it is a modeling artifact of variable role instantiation in the Dolev--Yao model.

\textbf{Important:} With \texttt{idp} as a lowercase constant, the intruder cannot forge
\texttt{idp}'s signature or impersonate the identity provider. The attack is purely about the
intruder playing a legitimate user role.

\textbf{Password security:} The guessable-password goal \texttt{pw(A,idp) guessable secret
between A, idp} is \emph{not} violated---this is the important Week~6 privacy result.

% ------------------------------------------------
\section{Assumptions, Simplifications, and Limitations}
\sectionauthor{Kristófer Birgir Hjörleifsson}

\subsection{Assumptions}
\begin{itemize}[leftmargin=*]
  \item \textbf{idp is honest:} Modeled as a lowercase constant in AnB. The intruder cannot
        instantiate the \texttt{idp} role.
  \item \textbf{Secure channels abstract TLS:} We assume TLS provides confidentiality and
        authentication as modeled by the channel notation.
  \item \textbf{Keys are pre-distributed:} Servers know each other's public keys (or obtain
        them via the key lookup subprotocol).
\end{itemize}

\subsection{Simplifications}
\begin{itemize}[leftmargin=*]
  \item \textbf{Photos as abstract secret:} Modeled as \texttt{photos(A)} rather than actual data.
  \item \textbf{No scope/folder:} Token authorizes access to all photos, not a specific subset.
  \item \textbf{No timestamps/nonces:} Token has no expiration or freshness mechanism.
  \item \textbf{Single session:} OFMC verifies a bounded number of sessions.
\end{itemize}

\subsection{Limitations}
\begin{itemize}[leftmargin=*]
  \item \textbf{Intruder can play A:} Since A, B, P are uppercase variables, the intruder can
        instantiate any of these roles. This leads to the photo secrecy attack where the
        intruder requests its own photos---a modeling artifact, not a real authorization break.
  \item \textbf{No revocation:} Once issued, tokens cannot be revoked.
  \item \textbf{idp single point of failure:} If \texttt{idp} were compromised (not modeled),
        all security guarantees fail.
  \item \textbf{No token freshness:} The token contains no nonce or timestamp, so replay
        of the token is possible within the OFMC bounded-session model.
\end{itemize}

% ------------------------------------------------
\section{Conclusion}
\sectionauthor{Mikael Máni Eyfeld Clarke}

We designed an open authorization protocol allowing user A to authorize printing service P to access
photos at server B, using identity provider \texttt{idp} for authentication.

\textbf{Key design decisions:}
\begin{itemize}[leftmargin=*]
  \item \texttt{idp} as lowercase constant prevents the intruder from impersonating the identity provider
  \item Format tags (\texttt{f1}--\texttt{f4}) prevent type-flaw attacks
  \item Pseudonymous channels (\texttt{[A] *->*}) model TLS with server-side authentication
  \item Hash-based challenge-response is vulnerable to offline guessing; secure channels are not
\end{itemize}

\textbf{Remaining attack:} OFMC reports ATTACK\_FOUND (secrets) because the intruder
can instantiate A with itself, causing \texttt{idp} to issue a legitimate token for the intruder's
own identity. The intruder then learns \texttt{photos(i)}---its own resource.
This is inherent to the Dolev--Yao model: any uppercase variable role can be played by the
intruder, and no secrecy goal formulation can prevent it. It is not a meaningful break of
honest-user authorization. The guessable-password goal holds, which is the key
Week~6 privacy result.

\textbf{Final protocol provides:}
\begin{itemize}[leftmargin=*]
  \item Authentication of \texttt{idp}'s authorization via signed tokens
  \item Confidentiality of photos between B and P (modulo the inherent Dolev--Yao modeling artifact)
  \item Protection of guessable password via pseudonymous TLS channels
\end{itemize}

% ------------------------------------------------
\appendix
\section{Submission Map (Where to Find Things)}
\sectionauthor{Kristófer Birgir Hjörleifsson}

\begin{itemize}[leftmargin=*]
  \item Lab logbook: \texttt{logbook/logbook.md}
  \item AnB protocol versions:
    \begin{itemize}
      \item \texttt{anb/week2\_v1.AnB} -- Basic protocol
      \item \texttt{anb/week3\_v1.AnB} -- Password authentication
      \item \texttt{anb/week4\_v1.AnB} -- Type-flaw resistant
      \item \texttt{anb/week5\_v1.AnB} -- Pseudonymous TLS channels
      \item \texttt{anb/week5\_v1\_tls.AnB} -- Crypto comparison (no channels)
      \item \texttt{anb/week6\_v1.AnB} -- Guessable password
      \item \texttt{anb/week6\_insecure.AnB} -- Insecure variant (demo)
    \end{itemize}
  \item Final AnB files:
    \begin{itemize}
      \item \texttt{anb/final/photo\_auth\_final.AnB} -- Final protocol
      \item \texttt{anb/final/key\_lookup.AnB} -- Key lookup subprotocol
    \end{itemize}
\end{itemize}

\end{document}